---
title: "nitrogen upshift of FY4 measured with coulter counter, analysis and figure generation"
author: "darach"
date: "`r Sys.Date()`"
---

The experiment:

- On 151130, one colony of DGY1 (FY4) yeast was started off a YPD
  plate into a 5ml O/N of NlimPro(800uM).
- On 151201, this was 1.17e7 cells per ml, so I diluted 400ul of this
  onto 50ml of NlimPro(800uM) @1120, put onto 30C shaker.
- On 151202, I took samples.

For each sample, I sampled into an eppendorf and sonicated them 
briefly (standard Gresham lab sonication for coulter counter), 
diluted them 1:50 in isoton (so 200ul into 10ml). These were measured
with Coulter Counter Z2. I measured a isoton
blank (recorded for below gating) then measured the sample, for each
one. I exported these as a Z2 file, and used the below script to
check the blanks for gating. We can't see below 2 micron diameter,
and so I set a cutoff of 2.5 microns because it cut out most of the
crud in the blanks.

During the experiment, I added 200ul 100mM glutamine at 14:09. 

```{r, libs, cache=F}
require(plyr)
require(reshape2)
require(ggplot2)
```

```{r, readinData, cache=T}
dataDir <- "../data/dme152coulterCounterUpshift/"
sampleSheet <- read.csv(paste0(dataDir,"SampleSheet.csv"),row.names=1)
sampleSheet$Minutes <- 
  60*as.numeric(sub("(\\d+?)(\\d\\d)","\\1",sampleSheet$Time))+
  as.numeric(sub("(\\d+?)(\\d\\d)","\\2",sampleSheet$Time))

datz <- ddply(
  ldply(list.files(path=dataDir,pattern=".=#Z2$"),.id="File",
    function(x) { print(paste0("Reading ",dataDir,"",x))
	    tmp <- readLines(paste0(dataDir,x))
	    binunits <- as.numeric(tmp[(grep("Bindiam",tmp)+1):
                                 (grep("Binunits",tmp)-1)])
	    binheight<- as.numeric(tmp[(grep("Binheight",tmp)+1):
                                 (grep("\\[end\\]",tmp)-1)])
	    return(data.frame(File=x,BinCenter=binunits,BinHeight=binheight))
    }),
  .(File),function(x){
    return(data.frame(x,sampleSheet[as.character(unique(x$File)),],
                      row.names=NULL))
  })
datz$Hours <- datz$Minutes/60
datz$Hours <- datz$Hours - (14+9/60)

datz[sample(1:nrow(datz),10),]
dim(datz)

```

```{r,exploratoryPlots, cache=T}

ggplot(datz)+aes(x=BinCenter,y=BinHeight,col=factor(Minutes))+
	geom_point()+facet_wrap(~Sample)

ggplot(subset(datz,BinCenter>2.0))+aes(x=BinCenter,y=BinHeight,col=factor(Minutes))+
	geom_point()+facet_wrap(~Sample)

ggplot(subset(datz,BinCenter>2.5))+aes(x=BinCenter,y=BinHeight,col=factor(Minutes))+
	geom_point()+facet_wrap(~Sample)

ggplot(subset(datz,BinCenter>3.0))+aes(x=BinCenter,y=BinHeight,col=factor(Minutes))+
	geom_point()+facet_wrap(~Sample)

datz <- subset(datz,BinCenter>2.5&BinCenter<7.0&Sample=="sample")
```

Upshift happened at 14:09, so that's at 849 minutes or 14.15 hours.

```{r, plots,cache=F}

counts <- aggregate(BinHeight~Hours,data=datz,FUN=sum)
counts$e6CellsPerMl <- counts$BinHeight/2000

# using log10 to make plotting behave nicely
firstGrowthModel <- lm(data=subset(counts,Hours<(0+2)),
  log10(e6CellsPerMl)~Hours)
prolineGR <- log(2)/firstGrowthModel$coefficients[2]/log(10)

secondGrowthModel <- lm(data=subset(counts,Hours>=(0+2)),
  log10(e6CellsPerMl)~Hours)
glutamineGR <- log(2)/secondGrowthModel$coefficients[2]/log(10)

g<-ggplot(counts)+aes(x=Hours,y=e6CellsPerMl)+theme_bw()+
  geom_point()+
  ylab("Million cells per ml\n")+
  xlab("Hours")+
  scale_y_log10(breaks=c(1,2,3,4,5,6,7,8,9,10,12,14))+
  scale_x_continuous(breaks=seq(-4,7,2))+
  annotate(geom="text",x=0.5,y=7,angle=270,
    label="Glutamine addition")+
  geom_vline(xintercept=0,linetype="solid")
g
g2<-g+geom_abline(intercept=firstGrowthModel$coefficients[1],
  slope=firstGrowthModel$coefficients[2],linetype="dotted",alpha=0.5)+
  annotate(geom="text",x=-2,y=4,
    label=paste0(signif(prolineGR,3)," hour\ndoubling time"))+
  geom_abline(intercept=secondGrowthModel$coefficients[1],
    slope=secondGrowthModel$coefficients[2],linetype="dotted",alpha=0.5)+
  annotate(geom="text",x=5,y=2.5,
    label=paste0(signif(glutamineGR,3)," hour\ndoubling time"))
g2
```

```{r,cache=T}
summaryStats <- ddply(datz,.(Hours),.id="Diameters",
  function(x){
    summary(rep(x$BinCenter,x$BinHeight))
  }
)

g <- ggplot(summaryStats)+theme_bw()+
	aes(x=Hours,y=Mean)+
	geom_point()+
  geom_vline(xintercept=0,linetype="solid")+
  geom_hline(yintercept=mean(subset(summaryStats,Hours<0)$Mean),
    linetype="dotted")+
  geom_hline(yintercept=mean(subset(summaryStats,Hours>2)$Mean),
    linetype="dotted")+
  scale_x_continuous(breaks=seq(-4,7,2))+
  ylab("Mean cell diameter,\nin microns")
g
```

```{r,cache=T}
mean(subset(summaryStats,Hours<0)$Mean)
mean(subset(summaryStats,Hours>2)$Mean)

g <- ggplot(melt(summaryStats,id.vars=c("Hours"),
              measure.vars=c("Median","Mean")))+
  theme_bw()+
	aes(x=Hours,y=value,col=variable)+facet_grid(variable~.)+
	geom_point()
g

together <- data.frame(counts,summaryStats)
together$log10e6CellsPerMl <- log10(together$e6CellsPerMl)

ggplot(melt(together,id.vars="Hours",
         measure.vars=c("log10e6CellsPerMl","Median")))+
  theme_bw()+
  aes(x=Hours,y=value)+facet_grid(variable~.,scale="free_y")+
  geom_point()+
  geom_vline(xintercept=0)+
  geom_vline(xintercept=2)

```

```{r}
library(tidyverse)
plotdatar <- full_join(counts%>%select(Hours,e6CellsPerMl),
  summaryStats%>%select(Hours,Median,Mean),
  by="Hours")%>%ungroup()
save(list=c("plotdatar","summaryStats"),file="../tmp/coulterCounterData.RData")
```

```{r}
sessionInfo()
```

