---
title: "Preliminary preliminary run of Quantigene probes, with a gap1 delete"
author: Darach
---


```{r}
require(tidyverse)
require(flowCore)
```

```{r}

datar <- list.files("../data/dme141",full.names=T) %>% 
  {tibble(Filename=.)} %>%
  group_by(Filename)%>%
  mutate(Data=list(as_tibble(
        exprs(read.FCS(Filename,transformation=F)))
      ))%>%unnest%>%
  mutate(Sample=c(
    `DAPI + APC -`="Negative (no probe)"
    ,`DAPI + APC +`="Proline-limited batch"
    ,`KO`="Negative (gap1 KO)"
    ,`Q`="Glutamine upshift, +20min")[
      sub(".*_([^_]+).fcs","\\1",Filename)
      ])%>%ungroup()%>%
  dplyr::select(-Filename)

range(datar$`APC-A`)
range(datar$`Pacific Blue-A`)

datar$`APC-A` <- datar$`APC-A` + 175
datar$`Pacific Blue-A` <-  datar$`Pacific Blue-A` +615

```

```{r, warning=F, cache=T}
tmp <- ggplot(datar[sample(1:nrow(datar)),])+theme_bw()
tmp+aes(x=log(`Pacific Blue-A`),y=log(`APC-A`),color=Sample)+geom_point(size=.5,alpha=.5)
```

And now, a histogram and then an approach that uses binning so that we can 
pull apart timepoints.

```{r, warning=F,cache=T}
hist(log(datar$`Pacific Blue-A`),100)
hist(log(datar$`APC-A`),100)

tmp+aes(x=log(`Pacific Blue-A`),col=factor(Sample))+
	stat_bin(aes(y = ..ncount..),binwidth=0.05,geom="line",position="identity",cex=.5)+
	stat_bin(aes(y = ..ncount..),binwidth=0.05,geom="point",position="identity",cex=2)+
    scale_x_continuous(breaks=seq(5,11,0.5))+
    theme(axis.text.x=element_text(angle=90))

tmp+aes(x=log(`APC-A`),col=factor(Sample))+
	stat_bin(aes(y = ..density..),binwidth=0.1,geom="line",position="identity",cex=.5)+
	stat_bin(aes(y = ..density..),binwidth=0.1,geom="point",position="identity",cex=2)+
    scale_x_continuous(breaks=seq(3,11,0.5))+
    theme(axis.text.x=element_text(angle=90))
```

Now, we plot things against each other. RNA will be on Y, DNA on X.

Before we try to get this to one dimensional plots, note that the DAPI signal
shifts with the gap1KO or the glutamine pulse. Strange. We need to use this
information in the gating.

```{r,cache=F,warning=F,error=F}

ggplot(subset(datar[sample(1:nrow(datar)),],
		Sample%in%c("Negative (no probe)","Proline-limited batch")&
		log(`Pacific Blue-A`)>7.5&log(`Pacific Blue-A`)<9.5)
	)+
	theme_bw()+
	aes(x=log(`Pacific Blue-A`),y=log(`APC-A`))+
	ylab("log Alexa647 (GAP1 mRNA)")+
	xlab("log DAPI (DNA)")+
	theme(legend.position="bottom")+
	geom_point(size=1.0,alpha=.1)+
	facet_wrap(~Sample)

ggplot(subset(datar[sample(1:nrow(datar)),],
		Sample%in%c("Negative (gap1 KO)","Glutamine upshift, +20min")&
		log(`Pacific Blue-A`)>7.25&log(`Pacific Blue-A`)<9.5)
	)+
	theme_bw()+
	aes(x=log(`Pacific Blue-A`),y=log(`APC-A`))+
	ylab("log Alexa647 (GAP1 mRNA)")+
	xlab("log DAPI (DNA)")+
	theme(legend.position="bottom")+
	geom_point(size=1.0,alpha=.1)+
	facet_wrap(~Sample)

subz <- subset(datar,
		(Sample%in%c("Negative (no probe)","Proline-limited batch")&
			log(`Pacific Blue-A`)>7.5&log(`Pacific Blue-A`)<9.5)
		|
		(Sample%in%c("Negative (gap1 KO)","Glutamine upshift, +20min")&
			log(`Pacific Blue-A`)>7.25&log(`Pacific Blue-A`)<9.5)
		)

ggplot(subz)+
	theme_bw()+
	aes(x=log(`Pacific Blue-A`),y=log(`APC-A`))+
	ylab("log Alexa647 (GAP1 mRNA)")+
	xlab("log DAPI (DNA)")+
	theme(legend.position="bottom")+
	geom_point(size=1.0,alpha=.1)+
	facet_wrap(~Sample)

```

Now that that's good, how does this look in one dimension?

```{r,cache=F,warning=F,error=F}
g <- subz %>% 
    ggplot()+theme_bw()+
	aes(x=log(`APC-A`),col=factor(Sample))+
	xlab("log Alexa647 (GAP1 mRNA)")+
	ylab("Count")+
	theme(legend.position="bottom")+
	stat_bin(aes(y = ..count..),binwidth=0.1,geom="line",position="identity",cex=.5)+
	stat_bin(aes(y = ..count..),binwidth=0.1,geom="point",position="identity",cex=2)+
    scale_color_discrete("")+
	xlim(c(3.5,10))
g
```

You'll note a few things. First, the population under induction 
looks very very bimodal. I believe this is a technical issue of
incomplete permeabilization, as it depends on fixation time. The 
later experiments were run with a time optimized on a poly-dT
probe, so I believe those results are much better. Do note that this
was much earlier, so the flourophore is probably fresher here!

Second, 20minutes after glutamine addition, it's still on? Nope, it
was likely repressed and re-induced. One of the super cool things
about this system is that GAP1 mRNA (and a bunch of other NCR)
are reinduced starting at around 15min, then they come back on.
We haven't even gotten past figuring out the first 10 minutes, so
we haven't really looked at that yet.

Anyways, for the supplemental figure I'm going to exclude the
glutamine re-induction one. 

```{r,cache=F,warning=F,error=F}
g <- subz %>% dplyr::filter(Sample!="Glutamine upshift, +20min")%>%
    ggplot()+theme_bw()+
	aes(x=log(`APC-A`),col=factor(Sample))+
	xlab(expression(paste("ln Alexa647 (",italic(GAP1)," mRNA)")))+
	ylab("Count")+
	theme(legend.position="bottom")+
	stat_bin(aes(y = ..count..),binwidth=0.1,geom="line",position="identity",cex=.5)+
	stat_bin(aes(y = ..count..),binwidth=0.1,geom="point",position="identity",cex=2)+
    scale_color_discrete("",labels=c(
      expression(paste(italic(gap1),Delta))
      ,expression(atop(italic(GAP1)*", no probe","negative control"))
      ,expression(atop(italic(GAP1)*", proline-limited","batch culture"))))+
	xlim(c(3.5,10))
g

ggsave("../output/Figure3_S_gap1deleteControl.un.tiff"
  ,g,width=5,height=5)

```
	

