---
title: "Procing and plotting dme209 facs data, figuring out the prior distributions..."
author: "Darach"
date: "`r Sys.Date()`"
---

# importing

First, libraries.

```{r}
library(tidyverse)
library(stringr)
```

No sample sheet, we just hardcode it.

```{r,readmeta,cache=T}
sampleSheet <- tibble(
  Filename=str_c("../data/dme209/facs/",
    c("Specimen_001_A1.fcs","Specimen_001_A4.fcs",
      "Specimen_001_B1.fcs","Specimen_001_B4.fcs",
      "Specimen_001_C1.fcs","Specimen_001_C4.fcs",
      "Specimen_001_I1-.fcs","Specimen_001_I1+.fcs",
      "Specimen_001_I2-.fcs","Specimen_001_I2+.fcs")
  ),
  BiologicalReplicate=c("A","A","B","B","C","C","I1","I1","I2","I2"),
  Shifted=c("pre-upshift","shifted")[c(1,2,1,2,1,2,1,1,1,1)],
  Probed=c("not-probed","probed")[c(2,2,2,2,2,2,1,2,1,2)]
)

```

Then we read the data in, and adjust up to scale everything into 
positive APC.A values.

```{r,readdatar,cache=T}
datar <- sampleSheet %>% group_by(Filename) %>%
  mutate(File=list(data.frame(flowCore::exprs(flowCore::read.FCS(Filename))))) %>% 
  unnest() %>% ungroup() 
gatez <- read_csv("../data/dme209/dme209.FACSGates.csv",comment="#") %>%
  mutate(Shifted=c("pre-upshift","shifted")[
    ifelse(FACSGate%in%c("p2","p3","p4","p5"),1,2)]) 

datar%>%filter(BiologicalReplicate%in%c("A","B","C"))%>%
  group_by(Shifted)%>%summarize(min(APC.A),max(APC.A))

2^18

datar%>%filter(BiologicalReplicate%in%c("A","B","C"))%>%
  group_by(Shifted)%>%summarize(min(Pacific.Blue.A),max(Pacific.Blue.A))

minApc <- -1200
minPb <- -150
datar <- datar %>% 
  mutate(APC.A=APC.A-minApc) %>% 
  mutate(Pacific.Blue.A=Pacific.Blue.A-minPb)
gatez <- gatez %>% 
  mutate(LowerBound=LowerBound-minApc,UpperBound=UpperBound-minApc)

gatez <- left_join(gatez,
  datar%>%
      filter(!(BiologicalReplicate%in%c("I1","I2")))%>%
      select(BiologicalReplicate,Shifted,APC.A) 
  ,by=c("BiologicalReplicate","Shifted")) %>%
  group_by(BiologicalReplicate,Shifted,FACSGate,LowerBound,UpperBound,Events) %>%
  summarize(MeanSignalInBin=
    mean(APC.A[APC.A>LowerBound&APC.A<=UpperBound+1],na.rm=T)) %>%
  ungroup()

```

```{r}
save(gatez,file="../tmp/dme209FACSgatez.RData")

```

```{r,exploreAndGate1,cache=T}
datar %>% ggplot(aes(x=FSC.A,y=FSC.A/FSC.W))+
  geom_point(size=0.1)+scale_y_log10()+scale_x_log10()+
  facet_grid(BiologicalReplicate~Shifted+Probed)
```


```{r,exploreAndGate2,cache=T}
datar %>% ggplot(aes(x=Pacific.Blue.A,y=APC.A))+
  geom_point(size=0.1,alpha=0.1)+scale_y_log10()+scale_x_log10()+
  facet_grid(BiologicalReplicate~Shifted+Probed)+
  geom_vline(xintercept=700,col="blue",linetype="dashed")
```

```{r}
flowDatar <- datar %>% filter(Pacific.Blue.A>700,BiologicalReplicate%in%c("A","B","C"))
save(flowDatar,file="../tmp/dme209flowCytometry.RData")
save(gatez,flowDatar,file="../tmp/Figure4facs.RData")
```

```{r,exploreAndGate3,cache=T}

g <- flowDatar %>% 
  ggplot(aes(x=APC.A))+theme_classic()+
    geom_histogram(aes(fill=Shifted),binwidth=0.05,position="identity",alpha=0.80)+
    guides(fill=F)+
    scale_x_log10(breaks=c(1,2,4,6,12,24)*1e3)+
    theme(axis.text.x=element_text(angle=90))+
    facet_grid(Shifted~BiologicalReplicate)+
    geom_vline(color="grey50",data=gatez,
      aes(xintercept=UpperBound))+
    geom_vline(color="grey50",data=gatez,
      aes(xintercept=LowerBound))+
#    geom_curve(linetype="solid",color="grey50",data=gatez,curvature=0,
#      arrow=arrow(ends="both",angle=90,length=unit(5,"inches")),
#      aes(x=LowerBound,xend=UpperBound,y=0,yend=0))+
    geom_text(data=gatez%>%group_by(FACSGate)%>%
      mutate(GateName=c("p2"="PreShift_1","p3"="Preshift_2",
        "p4"="Preshift_3","p5"="Preshift_4",
        "p6"="Shifted_1","p7"="Shifted_2",
        "p8"="Shifted_3","p9"="Shifted_4")[FACSGate],
        Center=mean(c(LowerBound,UpperBound))),
      aes(label=GateName,
      x=Center,
      y=220+(Shifted=="pre-upshift")*400),
      nudge_x=0.0,angle=90,size=3)+
    ylab("Events")+xlab("GAP1 mRNA FISH signal")
g+coord_cartesian(xlim=c(600,1e5))

gprezi <- flowDatar %>% 
  ggplot(aes(x=APC.A))+theme_classic()+
    geom_histogram(aes(fill=Shifted),binwidth=0.05,position="identity",alpha=0.80)+
    guides(fill=F)+
    scale_x_log10(breaks=c(1,2,4,6,12,24)*1e3)+
    theme(axis.text.x=element_text(angle=90))+
    facet_grid(Shifted~BiologicalReplicate)+
    geom_vline(color="grey50",data=gatez,
      aes(xintercept=UpperBound))+
    geom_vline(color="grey50",data=gatez,
      aes(xintercept=LowerBound))+
#    geom_curve(linetype="solid",color="grey50",data=gatez,curvature=0,
#      arrow=arrow(ends="both",angle=90,length=unit(5,"inches")),
#      aes(x=LowerBound,xend=UpperBound,y=0,yend=0))+
    ylab("Events")+xlab("GAP1 mRNA FISH signal")
gprezi+coord_cartesian(xlim=c(600,1e5))

```


```{r,modeling,cache=T}

g <- flowDatar %>% 
  ggplot(aes(x=log10(APC.A)))+theme_bw()+
    scale_x_continuous(breaks=0:9)+
    coord_cartesian()+#xlim=c(5e2,5e4))+
    geom_histogram(aes(fill=Shifted),bins=080,position="identity",alpha=0.80)+
    guides(fill=F)+
    theme(axis.text.x=element_text(angle=90))+
    facet_grid(Shifted~BiologicalReplicate)+
    ylab("Events")+xlab("GAP1 mRNA FISH signal")
g

g+stat_function(fun=function(x){dnorm(x,mean=3.15,sd=0.15)*3e2})

g2 <- g+coord_cartesian(xlim=log10(c(5e2,5e4)))

g2+stat_function(fun=function(x){dnorm(x,mean=3.15,sd=0.15)*3e2})
```



```{r,fittingDistributions,cache=T}

fitDatar <- flowDatar 

shifted <- data.frame(fitDatar %>% 
  filter(Shifted=="shifted") %>% 
  select(APC.A,BiologicalReplicate,Shifted))
shifted$APC.A <- log10(shifted$APC.A)

unshifted <- data.frame(fitDatar %>% 
  filter(Shifted=="pre-upshift") %>% 
  select(APC.A,BiologicalReplicate,Shifted))
unshifted$APC.A <- log10(unshifted$APC.A)

ggplot(rbind(shifted,unshifted),aes(x=APC.A,col=BiologicalReplicate))+
  facet_wrap(~Shifted)+geom_density()

```

Going to try to fit half a normal to the shifted populations.
I'm assuming that the right tail is real mutants, so I'm going to
reflect the left of mode one, and fit the sd to that.

```{r,fittingDistributions1,cache=T,warning=F}
shist <- hist(shifted$APC.A,150,plot=T)
leftmode <- shist$mids[which(shist$counts==max(shist$counts))]

reflectedLeft <- c(shifted$APC.A[shifted$APC.A<=leftmode],
  -(shifted$APC.A[shifted$APC.A<=leftmode]-leftmode)+leftmode)

leftshiftll <- function(sdz) {
  return(-sum(log(
    dnorm(reflectedLeft,mean=leftmode,sd=sdz)
  )))
}
leftmod <- stats4::mle(minuslogl=leftshiftll,start=list(sdz=1))

leftplot <- hist(shifted$APC.A,100,plot=F)
ggplot(data.frame(x=leftplot$mids,y=leftplot$density),aes(x=x,y=y))+
  geom_bar(stat="identity")+
  stat_function(size=2,alpha=0.5,col="blue",fun=function(x){
    dnorm(x,mean=leftmode,sd=leftmod@coef[1])*0.8})
```

Blue is the fit.  
Again, I think the right tail is bonafide mutants, or residual
repression from GAP1 not being completely off yet.

Either way, I'm looking for things that are suprisingly on,
compared to a distribution of mean `r leftmode` and sd 
`r leftmod@coef[1]`.

This is with a minApc of `r minApc`.

What about the right one?

```{r,fittingDistributions2cache=T,warning=F}

#unshifted <- subset(unshifted,BiologicalReplicate!="")

onll <- function(p1,meanz,sdz) {
  return(-sum(log(
    p1*dnorm(unshifted$APC.A,mean=leftmode,sd=leftmod@coef[1])+
    (1-p1)*dnorm(unshifted$APC.A,mean=meanz,sd=sdz)
  )))
}
onmod <- stats4::mle(minuslogl=onll,
  start=list(p1=0.1,meanz=3,sdz=.1))
onplot <- hist(unshifted$APC.A,100,plot=F)
ggplot(data.frame(x=onplot$mids,y=onplot$density),aes(x=x,y=y))+
  geom_bar(stat="identity")+
  stat_function(size=2,alpha=0.5,col="blue",fun=function(x){
    dnorm(x,mean=leftmode,sd=leftmod@coef[1])*onmod@coef[1]})+
  stat_function(size=2,alpha=0.5,col="yellow",fun=function(x){
    dnorm(x,mean=onmod@coef[2],sd=onmod@coef[3])*(1-onmod@coef[1])})+
  stat_function(size=2,alpha=0.5,col="green",fun=function(x){
    dnorm(x,mean=leftmode,sd=leftmod@coef[1])*onmod@coef[1]+
    dnorm(x,mean=onmod@coef[2],sd=onmod@coef[3])*(1-onmod@coef[1])})

```

So that's a mix of `r onmod@coef[1]` of an off normal of mean 
`r leftmode` and sd `r leftmod@coef[1]`, and a right normal of 
`r onmod@coef[2]` mean and `r onmod@coef[3]` sd.


```{r}
sessionInfo()
```

