---
title: "Modeling the pulse-chase results to identify a proper cutoff"
author: "Darach Miller"                                               
date: "Typeset `r Sys.Date()`"
---

For a more in depth analysis, see the Supplementary material.

```{r,libz,message=F,warning=F}
library(plyr)
library(tidyverse)
library(knitr)
library(stringr)
library(qvalue)
library(magrittr)
```

```{r,readingIn,cache=F}
load(file="../tmp/dme211datarForModeling.RData")
load(file="../tmp/dme211datarForModelingDirect.RData")
load(file="../tmp/dme211fitMaudzAltNorm.RData")
load(file="../tmp/dme211fitMaudz.RData")
load(file="../tmp/dme211normalizedData.RData")
load("../tmp/NameIDList.RData")
```

In the analysis detailed in the Supplementary material, we 
identified a cut-off for the rate changes that will better
select high-confidence destabilization events. Refer to that
document for the reasoning, this one is just to fit into the pipeline
nicely.

```{r}
bonafideAccelDegRates <- inner_join( maudzAltNorm%>%
    filter(Variable=="ChangeRate"&Estimate<0&QValue<0.01)%>%
    select(Systematic)
  ,maudzAltNorm%>%filter(Variable!="(Intercept)")%>%
  select(Systematic,Variable,Estimate)%>%
  spread(Variable,Estimate)%>%filter(BasalRate-ChangeRate>0)
  ,by="Systematic")
bonafideAccelDeg <- bonafideAccelDegRates%>%
  select(Systematic)%>%distinct()%>%unlist()
bonafideAccelDegRates%>%
  gather(Variable,Value,BasalRate,ChangeRate)%>%
  ggplot()+aes(x=Value)+geom_histogram()+facet_grid(Variable~.)

bonafideAccelDegRatesDirect <- inner_join( maudz%>%
    filter(Variable=="ChangeRate"&Estimate<0&QValue<0.05)%>%
    select(Systematic)
  ,maudz%>%filter(Variable!="(Intercept)")%>%
  select(Systematic,Variable,Estimate)%>%
  spread(Variable,Estimate)%>%filter(BasalRate-ChangeRate>0)
  ,by="Systematic")
bonafideAccelDegDirect <- bonafideAccelDegRatesDirect%>%
  select(Systematic)%>%distinct()%>%unlist()
bonafideAccelDegRatesDirect%>%
  gather(Variable,Value,BasalRate,ChangeRate)%>%
  ggplot()+aes(x=Value)+geom_histogram()+facet_grid(Variable~.)

tibble(Systematic=bonafideAccelDeg)%>%
  left_join(maudzAltNorm,by="Systematic")%>%
  filter(Variable!="(Intercept)")%>%
  select(Systematic,Variable,Estimate)%>%
  spread(Variable,Estimate)%>%distinct()
```

So, `r length(bonafideAccelDeg)` look accelerated more than
can be explained by in-complete label chase in the water samples,
and `r length(bonafideAccelDegDirect)` are in the direct 
normalization, with lower confidence.

```{r}
save(list=c("bonafideAccelDeg","bonafideAccelDegDirect")
  ,file="../tmp/dme211bonafideDegraders.RData")
```

```{r}
sessionInfo()
```
