---
title: "Doing PCA on the Airoldi et al 2016 microarray data"
author: "Darach"
date: "`r Sys.Date()`"
---

```{r,libs,cache=F}
library(pcaMethods)
library(tidyverse)
library(ggrepel)
library(stringr)
```

```{r, readindata,cache=T}
load("../tmp/airoldiEtAl2016MBoCTableS7.RData")
load("../tmp/NameIDList.RData")
ts7 <- ts7 %>% filter(!is.na(Systematic))
```

`ts7` is now our supplementary table from AiroldiEtAl2016, but we've
subset out the rows where Systematic name has been identified, which
is simply excluding some probes that are only present in the batch
upshift experiment. Different platform? Maybe also why it looks
like it starts from a slightly different place in the PCA.

Here we take out the data of steady-state growth at different rates
on different sources, and then the small glutamine pulses. 
There's a longterm, and a short timescale series. 
We're going to show the short term because although the long term
dynamics are super cool, we never followed up on this. Oh well.
This analysis is now just really to see if the batch is similar to
the chemostat, more than just changes in steady-state rate.

```{r,pca,cache=T}
ts7mat <- data.frame(ts7[,-c(1,2)])
rownames(ts7mat) <- as.character(ts7$Systematic)

smallQandSSandBatch <- ts7mat[,colnames(ts7mat)[
    grepl("(D\\.0\\.)|(40uM.Glutamine.*)|(Batch.400uM.Gln)",
    colnames(ts7mat))] ]
smallQandSSandBatch <- smallQandSSandBatch[complete.cases(smallQandSSandBatch),]
dim(smallQandSSandBatch)

sqsbPca <- pca(object=t(smallQandSSandBatch), method="svd", nPcs=30, 
  scale="uv", center=TRUE)
```

Below, we plot and subset out the components that explain more than
1% of the variance, and use those in the construction of an object
for plotting.

```{r,preplot,cache=T}

plot(log10(sqsbPca@R2));abline(h=log10(0.01))
ninetyNineScores <- scores(sqsbPca)[,sqsbPca@R2>0.01]
plotobj <- as_tibble(
  data.frame(SampleNames=rownames(ninetyNineScores),
    ninetyNineScores)) %>%
  mutate(SteadyState=grepl("D.0",SampleNames),
    DilutionRate=ifelse(grepl("D.0",SampleNames),
      sub(".*D\\.(0.\\d+)$","\\1",SampleNames),NA),
    PulseMinutes=ifelse(grepl("min$",SampleNames),
      sub(".*t\\.(\\d+)\\.min","\\1",SampleNames),NA),
    PulseSeconds=ifelse(grepl("seconds$",SampleNames),
      sub(".*t\\.(\\d+)\\.seconds","\\1",SampleNames),NA),
    Experiment="",Timescale=NA
  ) %>%
  mutate(Timescale=ifelse(grepl("(seconds$)|(Batch)",SampleNames),
    "Short-term","Long-term")) %>%
  mutate(Experiment=ifelse(grepl("400uM\\.G",SampleNames),
    "400uM pulse\nglutamine, ",Experiment)) %>%
  mutate(Experiment=ifelse(grepl("40uM\\.G",SampleNames),
    "40uM pulse\nglutamine, ",Experiment)) %>%
  mutate(Experiment=ifelse(grepl("^Batch.*min$",SampleNames),
    str_c(Experiment,"batch"),str_c(Experiment,"chemostat"))) %>%
  mutate(PulseMinutes=ifelse(grepl("seconds$",SampleNames),
      sub(".*t\\.(\\d+)\\.seconds$","\\1",SampleNames),PulseMinutes)) %>%
  mutate(PulseSeconds=ifelse(grepl("min$",SampleNames)
      ,as.numeric(sub(".*t\\.(\\d+)\\.min$","\\1",SampleNames))*60
      ,PulseSeconds)) %>%
  mutate(PulseMinutes=as.numeric(PulseMinutes),
    DilutionRate=as.numeric(DilutionRate)) %>%
  mutate(PulseMinutes=ifelse(grepl("seconds$",SampleNames),
      signif(PulseMinutes/60,3),PulseMinutes)) %>%
  arrange(PulseMinutes)

```


```{r,savePlotObj}
write_lines("#
# Supplementary Table
#
# Results of principal components analysis (using the pcaMethods 
# package) of Airoldi et al 2016 microarray data of steady-state 
# and dynamic increases in nitrogen-limited growth.
#
# Key:
# Name of sample , PC1 , PC2 , PC3 , PC4 , PC5 , PC6 , PC7 , PC8 , PC9 , PC10 , PC11 , PC12 , PC13 , PC14 , PC15 , is this sample at steady-state growth? , if chemostat, what is the dilution rate? , minutes after nitrogen addition , seconds after nitrogen addition , name of experiment for plotting , timescale of experiment" 
 ,path="../output/Figure1_Table_PCAresults.csv") 
plotobj %>% 
  write_csv(path="../output/Figure1_Table_PCAresults.csv",append=T)
```

Percent of variance explained ....

```{r}
sqsbPca@R2
```

```{r,plot,cache=F}

gss <- plotobj%>%unite(Experiment,Experiment,Timescale,sep="")%>%
  ggplot()+
  aes(col=DilutionRate)+
  theme_classic()+
  geom_point(data=plotobj%>%filter(SteadyState),size=2)+
  scale_color_gradient("Steady-state\ngrowth rate",
    low="#91bfdb",high="#fc8d59")+
  theme(legend.position="bottom",legend.key.size=unit(.5,"cm"))+
  guides(linetype=guide_legend(ncol=1))+
  scale_linetype_discrete("Upshift time-series"
    ,labels=c(`400uM pulse\nglutamine, batch`="batch"
      ,`40uM pulse\nglutamine, chemostat`="chemostat"
      )
    )
gss_pc12 <- gss+aes(x=PC1,y=PC2)+
  xlab(str_c("PC1\n",signif(sqsbPca@R2[1]*100,3),"% variance"))+
  ylab(str_c("PC2\n",signif(sqsbPca@R2[2]*100,3),"% variance"))+
  coord_cartesian(xlim=range(plotobj$PC1),ylim=range(plotobj$PC2))
gss_pc12

gall <- gss +
  geom_point(data=plotobj%>%filter(!SteadyState),size=2)+
  geom_text_repel(data=plotobj%>%filter(!SteadyState),
    box.padding=unit(0.5,"lines"),
    color="black",aes(label=paste(PulseSeconds,"s")))+
  geom_path(data=plotobj%>%filter(!SteadyState),
    aes(group=factor(Experiment):factor(Timescale),linetype=Experiment))
gall_pc12 <- gall+aes(x=PC1,y=PC2)+
  xlab(str_c("PC1\n",signif(sqsbPca@R2[1]*100,3),"% variance"))+
  ylab(str_c("PC2\n",signif(sqsbPca@R2[2]*100,3),"% variance"))+
  coord_cartesian(xlim=range(plotobj$PC1),ylim=range(plotobj$PC2))
gall_pc12

gtwo <- gss +
  geom_point(data=plotobj%>%filter(Timescale=="Short-term"),size=2)+
  geom_text_repel(data=plotobj%>%
    filter(!SteadyState,Timescale=="Short-term"),
    box.padding=unit(0.5,"lines"),
    color="black",aes(label=paste(PulseSeconds,"s"))
    )+
  geom_path(data=plotobj%>%
      filter(!SteadyState,Timescale=="Short-term")
    ,aes(group=factor(Experiment):factor(Timescale),linetype=Experiment))
gtwo_pc12 <- gtwo+aes(x=PC1,y=PC2)+
  xlab(str_c("PC1\n",signif(sqsbPca@R2[1]*100,3),"% variance"))+
  ylab(str_c("PC2\n",signif(sqsbPca@R2[2]*100,3),"% variance"))+
  coord_cartesian(xlim=range(plotobj$PC1),ylim=range(plotobj$PC2))
gtwo_pc12
```

What about higher PCs?

```{r,plotScores}
pcVsVariance <- data.frame(PC=names(sqsbPca@R2),FractionVariance=sqsbPca@R2)%>%
  mutate(PC=factor(PC,levels=PC))%>%
  ggplot()+aes(x=PC,y=FractionVariance)+geom_point()+
  theme_bw()+
  theme(axis.text.x=element_text(angle=90))+
  xlab("")

g34 <- gss +
  geom_point(data=plotobj%>%filter(Timescale=="Short-term"),size=2)+
  geom_text_repel(data=plotobj%>%
    filter(!SteadyState,Timescale=="Short-term"),
    box.padding=unit(0.5,"lines"),
    color="black",aes(label=paste(PulseSeconds,"s"))
    )+
  geom_path(data=plotobj%>%
      filter(!SteadyState,Timescale=="Short-term")
    ,aes(group=factor(Experiment):factor(Timescale),linetype=Experiment))+
  aes(x=PC3,y=PC4)+
  xlab(str_c("PC3\n",signif(sqsbPca@R2[3]*100,3),"% variance"))+
  ylab(str_c("PC4\n",signif(sqsbPca@R2[4]*100,3),"% variance"))+
  coord_cartesian(xlim=range(plotobj$PC3),ylim=range(plotobj$PC4))+
  theme(legend.position="right")

gHigher <- cowplot::plot_grid(pcVsVariance,g34,rel_widths=c(.4,.6),labels=c("A","B"))
ggsave("../output/Figure1_S_higherComponents.un.tiff",gHigher,width=7,height=5)

```

Next, we look for genes that correlate well with PC1 and PC2.

```{r,cor,cache=T}
pc1corz <- cor(scores(sqsbPca)[,1],t(smallQandSSandBatch))
pc2corz <- cor(scores(sqsbPca)[,2],t(smallQandSSandBatch))

hist(pc1corz)
hist(pc2corz)
```

And save everything.

```{r,outz}
save(plotobj,sqsbPca,pc1corz,pc2corz,file="../tmp/airoldiPCAanalysis.RData")
```

# Making the GO term tables.

```{r,clusterProfilerSetup,cache=T}
library(clusterProfiler)
library(org.Sc.sgd.db)
read_tsv("../data/sgd_go_slim_171013.txt",col_names=F                 
  ,comment="!") -> SGDGOSlim                                          
read_tsv("../data/sgd_go_terms_171013.txt",col_names=F                
  ,comment="!") %>%                                                   
  mutate(GOID=str_c("GO:"                                             
    ,str_pad(string=X1,width=7,side="left",pad="0"))                  
    ) -> SGDGOTerms 

pc1vector <- setNames(pc1corz[1,]
  ,nm=unlist(map(colnames(pc1corz),sysToSGD))
  )
pc2vector <- setNames(pc2corz[1,]
  ,nm=unlist(map(colnames(pc2corz),sysToSGD))
  )

nPermutations <- 1e6
GSEAResults <- list()
```

```{r,highPC1,dependson="clusterProfilerSetup",cache=T}
system.time(
GSEAResults[["PC1"]] <- GSEA(
  gene=sort(pc1vector,decreasing=T)
  ,TERM2GENE=SGDGOSlim%>%dplyr::select(X6,X3)
  ,TERM2NAME=SGDGOTerms%>%dplyr::select(GOID,X2)
  ,minGSSize=10,maxGSSize=500
  ,pAdjustMethod="fdr"
  ,pvalueCutoff=0.05,seed=171013
  ,nPerm=nPermutations
  )@result %>% mutate(PC="PC1")
)
```

```{r,highPC2,dependson="clusterProfilerSetup",cache=T}
system.time(
GSEAResults[["PC2"]] <- GSEA(
  gene=sort(pc2vector,decreasing=T)
  ,TERM2GENE=SGDGOSlim%>%dplyr::select(X6,X3)
  ,TERM2NAME=SGDGOTerms%>%dplyr::select(GOID,X2)
  ,minGSSize=10,maxGSSize=500
  ,pAdjustMethod="fdr"
  ,pvalueCutoff=0.05,seed=171013
  ,nPerm=nPermutations
  )@result %>% mutate(PC="PC2")
)
```

```{r,printOut,cache=T}

gseGOtable <- bind_rows(GSEAResults) %>% as_tibble %>% 
  dplyr::select(PC,ID,Description,enrichmentScore,p.adjust
    ,setSize,rank) %>%
  arrange(PC,-enrichmentScore)

write_lines("#
# Supplementary Table
#
# Results of gene set enrichment analysis of gene list ordered by 
# correlation to principal components 1 or 2 of the PCA of 
# Airoldi et al 2016 microarray data of steady-state and dynamic 
# increases in nitrogen-limited growth.
#
# Key:
# Principal component , GO term ID , GO term , enrichment score , adjusted p-value , size of GO term set , rank of this set in gene list"
  ,path="../output/Figure1_Table_GSEofGOtermsAgainstPCcorrelation.csv") 
gseGOtable %>% 
  write_csv(path="../output/Figure1_Table_GSEofGOtermsAgainstPCcorrelation.csv",append=T)

z <- gseGOtable %>% group_by(PC,enrichmentScore>0) %>%
  summarize(AllDescriptions=list(Description))  %>%
  spread(PC,AllDescriptions) %>%
  rowwise() %>%
  mutate(
    diffPC1=list(setdiff(unlist(PC1),intersect(unlist(PC1),unlist(PC2))))
    ,diffPC2=list(setdiff(unlist(PC2),intersect(unlist(PC1),unlist(PC2))))
    ,samePC1=list(intersect(unlist(PC1),unlist(PC2)))
    ,samePC2=list(intersect(unlist(PC1),unlist(PC2)))
    ) 
z
```

First, the ones that are shared between components. Remember, row 2 is
increased, 1 is decreased.

```{r,print2,cache=T}
z[2,]%>%pull(samePC1)
z[2,]%>%pull(samePC2)
z[1,]%>%pull(samePC1)
z[1,]%>%pull(samePC2)
```

Then, the ones that are distinct.
Remember, row 2 is increased, 1 is decreased.

```{r,print3,cache=T}
z[2,]%>%pull(diffPC1)
z[1,]%>%pull(diffPC1)

z[2,]%>%pull(diffPC2)
z[1,]%>%pull(diffPC2)
```

```{r}
sessionInfo()
```

